Simulation connectance richness

### 2023-02-28 report/cs_analysis.Rmd
Add Z to report

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::psim_document2
---

@@

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)

@@

theme_set(mytheme)
```
```{r}
sim_brut <- open_dataset(here::here("res", "sim_cs_stdK.arrow"),
  format = "arrow")
sim <- sim_brut %>%
  collect() %>%
  mutate(across(where(is.list), as.list)) %>%
  mutate(across(c(int_strength, max_int),
      ~map2(.x, S,
        function(y, rich) {
          matrix(y, nrow = rich)
        }
      )
  )
    ) %>%
  mutate(max_int_alive  = map2(max_int, bm_sp,
      function(mat, bm) {
        mask <- bm > 10^-16
        mat[mask, mask]
      }
      ),
    ct_alive = map_dbl(max_int_alive, ~ sum(.x > 0) / ((ncol(.x) - 1)^2)),
    bm_sp_alive = map(bm_sp, ~.x[.x > 10^-16]),
    tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-16]),
    max_tlvl_alive = map_dbl(tlvl_alive, max),
    w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive, ~sum(.x * (.y / sum(.y)))),
    max_max_int_alive = map_dbl(max_int_alive, ~max(.x[.x > 0])),
    avg_max_int_alive = map_dbl(max_int_alive, ~mean(.x[.x > 0])),
    #gini_max_int_alive = map_dbl(max_int_alive, ),
  )
```
```{r}
simZ <- open_dataset(here::here("res", "sim_csZ.arrow"),
  format = "arrow") %>%
  collect() %>%
  filter(!is.na(richness)) %>%
  mutate(across(where(is.list), as.list)) %>%
  mutate(across(c(int_strength, max_int),
      ~map2(.x, S,
        function(y, rich) {
          matrix(y, nrow = rich)
        }
      )
  )
    ) %>%
  mutate(max_int_alive  = map2(max_int, bm_sp,
      function(mat, bm) {
        mask <- bm > 10^-16
        mat[mask, mask]
      }
      ),
    ct_alive = map_dbl(max_int_alive, ~ sum(.x > 0) / ((ncol(.x) - 1)^2)),
    bm_sp_alive = map(bm_sp, ~.x[.x > 10^-16]),
    tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-16]),
    max_tlvl_alive = map_dbl(tlvl_alive, max),
    w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive, ~sum(.x * (.y / sum(.y)))),
    max_max_int_alive = map_dbl(max_int_alive, ~max(.x[.x > 0])),
    avg_max_int_alive = map_dbl(max_int_alive, ~mean(.x[.x > 0])),
    #gini_max_int_alive = map_dbl(max_int_alive, ),
  )
```

@@

```{r}
library(piecewiseSEM)
library(semEff)
```

```{r}
sim_sem <- sim %>%
  filter(rho == 0, env_stoch == .5) %>%
  mutate(across(c(sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}"))
```

```{r}
pairs(
  sim_sem %>%
    select(ct_alive, log_richness, max_tlvl_alive,
      log_gini_max_int, log_avg_max_int_alive, log_sync,
      log_avg_cv_sp, log_stab_com),
    pch = 19)
```

@@

```{r}
colnames(simZ)
corrplot::corrplot(
  cor(simZ %>%
    mutate(persistence = richness / S) %>%
    select(
      c(S, richness, persistence, ct, ct_alive, total_biomass, avg_max_int_alive,
        max_max_int_alive, gini_max_int, w_avg_tlvl_alive, max_tlvl_alive, Z,
        rho, env_stoch, stab_com, avg_cv_sp, sync
        )
      ),
    method = "spearman"),
  diag = FALSE, type = "upper"
)
```

>>>>

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::pdf_document2
---

@@

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)

@@

theme_set(mytheme)
```
```{r}
sim_brut <- open_dataset(
  here::here("res", "sim_cs.arrow"),
  #here::here("res", "sim_cs_stdK.arrow"),
  format = "arrow")
sim <- sim_brut %>%
  collect() %>%
  mutate(across(where(is.list), as.list)) %>%
  mutate(across(c(int_strength, max_int),
      ~map2(.x, S,
        function(y, rich) {
          matrix(y, nrow = rich)
        }
      )
  )
    ) %>%
  mutate(max_int_alive  = map2(max_int, bm_sp,
      function(mat, bm) {
        mask <- bm > 10^-16
        mat[mask, mask]
      }
      ),
    ct_alive = map_dbl(max_int_alive, ~ sum(.x > 0) / ((ncol(.x) - 1)^2)),
    bm_sp_alive = map(bm_sp, ~.x[.x > 10^-16]),
    tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-16]),
    max_tlvl_alive = map_dbl(tlvl_alive, max),
    w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive, ~sum(.x * (.y / sum(.y)))),
    max_max_int_alive = map_dbl(max_int_alive, ~max(.x[.x > 0])),
    avg_max_int_alive = map_dbl(max_int_alive, ~mean(.x[.x > 0])),
    #gini_max_int_alive = map_dbl(max_int_alive, ),
  )
```
```{r}
simZ <- open_dataset(here::here("res", "sim_csZ.arrow"),
  format = "arrow") %>%
  collect()
simZ2 <- open_dataset(here::here("res", "sim_csZ2.arrow"),
  format = "arrow") %>%
  collect()

# TODO set extinction threshold in simulations
simZ <- map_dfr(list(simZ, simZ2),
  function (x) {
    x %>%
      filter(!is.na(richness)) %>%
      mutate(across(where(is.list), as.list)) %>%
      mutate(across(c(int_strength, max_int),
          ~map2(.x, S,
            function(y, rich) {
              matrix(y, nrow = rich)
            }
          )
      )
        ) %>%
      mutate(max_int_alive  = map2(max_int, bm_sp,
          function(mat, bm) {
            mask <- bm > 10^-16
            mat[mask, mask]
          }
          ),
        ct_alive = map_dbl(max_int_alive, ~ sum(.x > 0) / ((ncol(.x) - 1)^2)),
        bm_sp_alive = map(bm_sp, ~.x[.x > 10^-16]),
        tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-16]),
        max_tlvl_alive = map_dbl(tlvl_alive, max),
        w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive, ~sum(.x * (.y / sum(.y)))),
        max_max_int_alive = map_dbl(max_int_alive, ~max(.x[.x > 0])),
        avg_max_int_alive = map_dbl(max_int_alive, ~mean(.x[.x > 0])),
        #gini_max_int_alive = map_dbl(max_int_alive, ),
      )

  }
) %>%
filter(max_max_int_alive != -Inf) %>%
mutate(persistence = richness / S)
```

@@

```{r}
sim_sem <- sim %>%
  filter(rho == 0, env_stoch == .5) %>%
  mutate(across(c(sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  as.data.frame() %>%
  select(c(ct_alive, log_richness, max_tlvl_alive, Z, gini_max_int, log_avg_max_int_alive, env_stoch, log_stab_com, log_sync, log_avg_cv_sp))
```

```{r}
pairs(
  sim_sem %>%
    select(ct_alive, log_richness, max_tlvl_alive,
      gini_max_int, log_avg_max_int_alive, log_sync,
      log_avg_cv_sp, log_stab_com),
    pch = 19)
```

@@

```{r}
colnames(simZ)
corrplot::corrplot(
  cor(simZ %>%
    select(
      c(S, richness, persistence, ct, ct_alive, total_biomass, avg_max_int_alive,
        max_max_int_alive, gini_max_int, w_avg_tlvl_alive, max_tlvl_alive, Z,
        rho, env_stoch, stab_com, avg_cv_sp, sync
        )
      ),
    method = "spearman"),
  diag = FALSE, type = "upper"
)
```

```{r}
unique(simZ$richness)
simZ %>%
  ggplot(aes(x = as.factor(Z), y = avg_cv_sp, fill = richness)) +
  geom_boxplot()
```

```{r}
sim_sem <- simZ %>%
  filter(S == richness) %>%
  filter(rho == 0) %>%
  mutate(across(c(sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  select(c(ct_alive, log_richness, max_tlvl_alive, Z, gini_max_int, log_avg_max_int_alive, env_stoch, log_stab_com, log_sync, log_avg_cv_sp)) %>%
  as.data.frame()
```

```{r}
sem_int_list <- list(
  ct = lm(ct_alive ~ log_richness, sim_sem),
  max_tlvl = lm(max_tlvl_alive ~ log_richness + Z + ct_alive, sim_sem),
  gini_max_int = lm(gini_max_int ~ log_richness + ct_alive + max_tlvl_alive + Z, sim_sem),
  log_avg_max_int = lm(log_avg_max_int_alive ~ log_richness + ct_alive + max_tlvl_alive + Z, sim_sem),
  log_sync = lm(log_sync ~ log_richness + log_avg_max_int_alive + max_tlvl_alive + gini_max_int + env_stoch + Z, sim_sem),
  log_avg_cv_sp = lm(log_avg_cv_sp ~ log_richness + log_avg_max_int_alive + max_tlvl_alive + gini_max_int + env_stoch + Z, sim_sem),
  log_stab_com = lm(log_stab_com ~ log_sync + log_avg_cv_sp, sim_sem)
)

```{r}
summary(sim_sem$richness)
summary(sim_sem$S)
summary(sim_sem$persistence)
```

```{r}
boot_int <- semEff(sem_int, ci.type = "perc", R = 50, ncpus = 2)
summary(boot_int)
```

```{r}
unique(simZ$Z)
sim_sem <- simZ %>%
  filter(rho == 0, Z == 50) %>%
  mutate(across(c(sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  select(c(ct_alive, log_richness, max_tlvl_alive, Z, gini_max_int, log_avg_max_int_alive, env_stoch, log_stab_com, log_sync, log_avg_cv_sp)) %>%
  as.data.frame()
```

```{r}
sem_int_list <- list(
  ct = lm(ct_alive ~ log_richness, sim_sem),
  max_tlvl = lm(max_tlvl_alive ~ log_richness + ct_alive, sim_sem),
  gini_max_int = lm(gini_max_int ~ log_richness + ct_alive + max_tlvl_alive , sim_sem),
  log_avg_max_int = lm(log_avg_max_int_alive ~ log_richness + ct_alive + max_tlvl_alive, sim_sem),
  log_sync = lm(log_sync ~ log_richness + log_avg_max_int_alive + max_tlvl_alive + gini_max_int + env_stoch , sim_sem),
  log_avg_cv_sp = lm(log_avg_cv_sp ~ log_richness + log_avg_max_int_alive + max_tlvl_alive + gini_max_int + env_stoch , sim_sem),
  log_stab_com = lm(log_stab_com ~ log_sync + log_avg_cv_sp, sim_sem)
)

```{r}
boot_int <- semEff(sem_int, ci.type = "perc", R = 50, ncpus = 2)
summary(boot_int)
```

```{r}
unique(simZ$Z)
sim_sem <- simZ %>%
  filter(rho == 1) %>%
  mutate(across(c(ct_alive, sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, .names = "log_{.col}")) %>%
  select(c(log_ct_alive, log_richness, max_tlvl_alive, Z, gini_max_int, log_avg_max_int_alive, env_stoch, log_stab_com, log_sync, log_avg_cv_sp)) %>%
  as.data.frame()
pairs(sim_sem, pch = 19)
```

```{r}
sem_int_list <- list(
  ct = lm(log_ct_alive ~ log_richness, sim_sem),
  max_tlvl = lm(max_tlvl_alive ~ log_richness + Z, sim_sem),
  gini_max_int = lm(gini_max_int ~ log_ct_alive + max_tlvl_alive , sim_sem),
  log_avg_max_int = lm(log_avg_max_int_alive ~ log_ct_alive + max_tlvl_alive + Z, sim_sem),
  log_sync = lm(log_sync ~ log_richness + log_avg_max_int_alive + gini_max_int + env_stoch, sim_sem),
  log_avg_cv_sp = lm(log_avg_cv_sp ~ log_richness + log_avg_max_int_alive + gini_max_int + env_stoch , sim_sem),
  log_stab_com = lm(log_stab_com ~ log_sync + log_avg_cv_sp, sim_sem)
)

```{r}
boot_int <- semEff(sem_int, ci.type = "perc", R = 50, ncpus = 2)
summary(boot_int)
```

```{r}
simZ %>%
  ggplot(aes(y = persistence, x = as.factor(Z))) +
  geom_boxplot()
```
